---
title: "Temperature responses"
author: "Til H채mmig, Francesco Polazzo, Owen L. Petchey, Frank Pennekamp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
    keep_md: no
    fig_caption: false  
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
rm(list=ls())
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("here")
library(flextable)
library(performance)
library(lavaan)
library(officer)
library(gt)
library(ggbeeswarm)
library(lmerTest)
library(codyn)
library(broom)

#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")

```

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```
# Introduction

Central to the concept of response diversity is that species respond to a change in the environment, in our case a fluctuation in temperature. Our monoculture data clearly showed that species performances changed along the temperature gradient. Here we provide further evidence that this holds true in the community experiment. We relate species responses in monoculture to the regression coefficient of temperature against biomass, and find a positive relationship, meaning that positive responses were associated with an increase in biomass with temperature while negative responses were related to a decrease (Fig. 1 & 2).




```{r  warning=FALSE, results='hide'}

divergence_df <- read_csv(here("Data", "divergence_df.csv"))
load(here("Data", "dens_biomass_poly.RData"))

dd_all_pred<-read.csv(here("Data", "morph_dd_pred.csv"))
dd_all_pred_nonoise<-read.csv(here("Data", "morph_dd_pred_nonoise.csv"))

load(here("Data", "ciliate_traits.Rdata"))

df_slopes <- read_csv(here("Data", "df_slopes_cor.csv"))

# needs to have id_new variable
ciliate_traits <- ciliate_traits %>%
  dplyr::mutate(
    # Remove dots from the date
    cleaned_date = gsub("\\.", "", date),
    # Extract the part of id after the underscore
    id_suffix = sub(".*_(.*)", "\\1", id),
    # Combine cleaned_date, id_suffix, and species_initial into a new variable
    id_new = paste0(cleaned_date, id_suffix, composition)
  ) %>%
  # Optionally, remove the intermediate columns to clean up
  dplyr::select(-cleaned_date, -id_suffix,-new_id)

uniqueN(ciliate_traits$id_new)==nrow(ciliate_traits) # all unique  ;)

id_dd<-full_join(dd_all_pred,dplyr::select(ciliate_traits,id_new,biomass),join_by("id_new"))


## add day variable

#create a day variable from the date variable

id_dd<-dplyr::mutate(id_dd,date=as.Date(date,format = "%d.%m.%y"))

earliest_date<-min(id_dd$date)
days_since_earliest<-as.numeric(id_dd$date-earliest_date)+1
id_dd<-id_dd%>%dplyr::mutate(day=days_since_earliest)

#create a summarised df on microcosm level with each species seperate
# Make sure, that we have n_frames and not N_frames
names(id_dd)[names(id_dd) == "N_frames"] <- "n_frames"

#extrapolation_factor <- 9.301902  # for 16 x magnification 
extrapolation_factor <- 9.828125  # for 25 x magnification 
video_biomass_species <- c( "C", "P", "S","D","L","T")

biomasses <- id_dd %>%
  dplyr::group_by( day,temperature,nutrients,sample_ID,composition,predict_spec) %>% # group  by xxx
  dplyr::summarize(
    biomass = sum(biomass * n_frames, na.rm = TRUE) / (1 * 125) # if not 3 videos corrections is done below with dens_factor
  ) %>%
  dplyr::mutate(
    biomass = biomass * extrapolation_factor,
    )

biomasses<-biomasses%>%dplyr::mutate(biomass=biomass*1000)

dd_ts_id<-biomasses

#fill up missing dates with biomass<-0

fill_dd<-expand.grid(sample_ID=unique(dd_ts_id$sample_ID),day=unique(dd_ts_id$day),predict_spec=unique(dd_ts_id$predict_spec))
complete_ts<-full_join(fill_dd,dd_ts_id,join_by(sample_ID,day,predict_spec))


complete_ts$biomass[is.na(complete_ts$biomass)]<-0
complete_ts<-complete_ts%>%dplyr::mutate(composition=sub("_.*", "", sample_ID))
complete_ts<-complete_ts %>%
  dplyr::mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
complete_ts<- dplyr::mutate(complete_ts,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))

# Now remove wrong combinations of composition and predict_spec / predict_spec

complete_ts<- complete_ts %>%
  rowwise() %>%
  dplyr::filter(predict_spec %in% unlist(strsplit(composition, ""))) %>%
  ungroup()  
complete_ts<-dplyr::mutate(complete_ts,temperature=as.character(temperature),
                    nutrients=as.character(nutrients),
                    richness=nchar(composition))

complete_ts<-complete_ts%>%group_by(sample_ID,composition,day)%>%dplyr::mutate(tot_biomass=sum(biomass))
complete_ts<-complete_ts%>%dplyr::mutate(biom_contribution=biomass/tot_biomass)

df_biomass_mod <- complete_ts

complete_ts<-complete_ts%>%dplyr::mutate(temperature=paste0(temperature," 째C"),
                                      nutrients=paste0(nutrients," g/L"))


# introduce slopes  
names(df_slopes)[names(df_slopes)=="species_initial"]<-"predict_spec"

slope_ts<-full_join(dplyr::select(df_slopes,nutrients,predict_spec,temperature,slope),complete_ts)
slope_ts<-slope_ts%>%dplyr::mutate(w_slope=biom_contribution*slope,
                            sign=sign(slope))

slope_ts<-slope_ts%>%group_by(sample_ID,temperature,nutrients,richness,composition,day,tot_biomass)%>%dplyr::summarize(
  sum_w_slopes=abs(sum(w_slope)),
                   mean_abs_slope=mean(abs(slope)),
  sum_abs_slope=sum(abs(slope)),
  abs_sum_slope=abs(sum(slope)),
  symmetry=abs(sum(sign)))


slope_ts<-slope_ts%>%dplyr::mutate(richness=as.factor(richness))


##create new variable where it checks, where the last observation =0 is; with complete_ts
aggr_ts <- slope_ts %>%
  group_by( sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = tot_biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()

aggr_ts<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))







```


 
## Species responses to temperature

  T0:    Does temperature on sampling day explain biomass?  
  
  T1:   Does temperature on the day before sampling day explain biomass? 
  
  T2:   Does average temperature over the two days before sampling day explain biomass? 
  
  T3:   Does average temperature over the three days before sampling day explain biomass? 



```{r}

## Create temperature time series patterns
species_ts<-complete_ts%>%mutate(species_ID=paste(sample_ID,predict_spec,sep="_"))

pattern <- c(18, 18, 18, 19.5, 21, 21, 21, 19.5)
T1<-rep(pattern, length.out = 60)
T1_sel<-data.frame(sample_T=T1[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T1, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T1, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T1, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))


pattern <- c(22, 22, 22, 23.5, 25, 25, 25, 23.5)
T2<-rep(pattern, length.out = 60)
T2_sel<-data.frame(sample_T=T2[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T2, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T2, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T2, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))

pattern <- c(25, 25, 25, 26.5, 28, 28, 28, 26.5)
T3<-rep(pattern, length.out = 60)
T3_sel<-data.frame(sample_T=T3[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T3, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T3, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T3, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))


## Note that total time series goes from sampling day 1 (second day of the experiment) until the 59th day (here 58)-> 1:58

library(dplyr)
library(slider)

# Create the temperature time series for each regime


# Get unique days 
days <- unique(species_ts$day)+1 

# Helper function to compute sliding means
get_temp_summary <- function(temp_vector, days) {
  tibble(
    day = days-1 ,  # shift back to match original day indexing
    mean_T_lag1 = sapply(days, function(d) if (d >= 2) mean(temp_vector[(d-1)]) else NA)%>%unlist,
    mean_T_lag2 = sapply(days, function(d) if (d >= 3) mean(temp_vector[(d-2):(d-1)]) else NA)%>%unlist,
    mean_T_lag3 = sapply(days, function(d) if (d >= 4) mean(temp_vector[(d-3):(d-2)]) else NA)%>%unlist
  )
}

# Generate summary tables
T1_mean_lag <- get_temp_summary(T1, days)
T2_mean_lag <- get_temp_summary(T2, days)
T3_mean_lag <- get_temp_summary(T3, days)


T1_sel<-full_join(T1_sel,T1_mean_lag)
T2_sel<-full_join(T2_sel,T2_mean_lag)
T3_sel<-full_join(T3_sel,T3_mean_lag)



df_S<-split(species_ts,species_ts$species_ID)
df_S<-mclapply(df_S,function(df){
  if(unique(df$temperature)=="18-21 째C"){
    dfT<-T1_sel
  }
  else if(unique(df$temperature)=="22-25 째C"){
    dfT<-T2_sel
  }
  else{dfT<-T3_sel 
  }
  

  
   dfT<-full_join(dfT,df)
   
   dfT<-dfT %>%
  group_by( species_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()
   

  
   
  
  })
   df_T<-do.call("rbind",df_S)



list_reg<-lapply(df_S,function(s){
  # Filter data
  s_filtered <- s %>% dplyr::filter(is_after_first_zero_day == F,day<=30)
    
  
  if (nrow(s_filtered) < 3) {
    return(NULL)
    }
  
  



  # Fit models
  lm0 <- lm((biomass) ~ sample_T, data = s_filtered)
  lm1 <- lm((biomass) ~ mean_T_lag1, data = s_filtered)
  lm2 <- lm((biomass) ~ mean_T_lag2, data = s_filtered)
  lm3 <- lm((biomass) ~ mean_T_lag3, data = s_filtered)
 
  # Get species responses
  slope_select<-df_slopes%>%dplyr::filter(nutrients==unique(s$nutrients),temperature==unique(s$temperature),predict_spec==unique(s$predict_spec))

 
  
  data.frame(
    species_ID = unique(s$species_ID),
    coef_sample_T = coef(lm0)[["sample_T"]],
    coef_temp_lag1 = coef(lm1)[["mean_T_lag1"]],
    coef_temp_lag2 = coef(lm2)[["mean_T_lag2"]],
    coef_temp_lag3 = coef(lm3)[["mean_T_lag3"]],
    slope=(slope_select$slope),
    temperature=unique(s$temperature),
    nutrients=unique(s$nutrients),
    predict_spec=unique(s$predict_spec),
    mean_biom_cont=mean(s_filtered$biom_contribution,na.rm=T),
    richness=nchar(unique(s_filtered$composition)),
    last_non_zero_day=unique(s$last_non_zero_day)
    )
  
})

df_reg<-do.call(rbind,list_reg)



temp_lags <- c("coef_sample_T", "coef_temp_lag1", "coef_temp_lag2", "coef_temp_lag3")

# Apply cor.test for each temp lag and tidy the results
cor_summary <- map_dfr(temp_lags, function(var) {

  result <- cor.test(df_reg$slope, df_reg[[var]])
  broom::tidy(result) %>%
    mutate(temp_lag = var) %>%
    select(temp_lag, estimate, conf.low, conf.high, p.value)%>%
    mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2"  ~"T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))
})


plot_overal_cor<-ggplot(data = cor_summary ,
       aes(x = temp_lag%>%as.factor, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +

  geom_hline(yintercept=0,linetype="dashed")+
  theme_bw() +
  scale_x_discrete(labels = function(x) parse(text = x)) +
  labs( x = "Temperature variable",y="Correlation")




df_reg_long <- df_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2"  ~"T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))






```



## Overall correlation


```{r}
df_reg_long <- df_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2" ~ "T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))




plot_cor_reg_slope <- ggplot(data = df_reg_long, aes(x = slope, y = coef_value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~temp_lag, nrow = 1) +
  theme_bw(base_size = 20)+
  labs(y="Regression coefficient: biomass ~ temperature",x = "Response from monoculture (slope)") 
```



```{r overal reg_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_cor_reg_slope
```



**Figure 1**: Relationship between species' responses to temperature from monoculture and from the community experiment. The positive relationship is evidence that species were responding to the fluctuating temperature regime in the community experiment. 

```{r overal cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
 plot_overal_cor
```



**Figure 2**: Correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope). Positive correlations indicate that species in polyculture were responding according to the responses estimated from the monocultre experiment. 




## Species level

Depending on the magnitude of the different responses more time is required for a change in performance to result in a detectable change in biomass. We looked at species individually to understand how responsive they were to temperature fluctuations (Tab. 1 & Fig. 3). We found strong evidence for Colpidium (C), and weak / mixed evidence for all the other species. Colpidium had by far the strongest responses (largest slopes) (Fig. 4). Thus, it is expected that we would also find the clearest signal. 

```{r species cor_table, results='asis', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
# Prepare correlation results
cor_table <- df_reg %>%
  pivot_longer(cols = starts_with("coef_"), names_to = "coef_var", values_to = "coef_val") %>%
  group_by(predict_spec, coef_var) %>%
  summarise(cor_res = list(cor.test(slope, coef_val) %>% tidy(conf.int = TRUE)), .groups = "drop") %>%
  unnest(cor_res) %>%
  select(predict_spec, coef_var, estimate, conf.low, conf.high, p.value) %>%
  mutate(
    estimate = round(estimate, 3),
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3),
    p.value = signif(p.value, 2),
    `95% CI` = paste0("[", conf.low, ", ", conf.high, "]")
  ) %>%
  select(predict_spec, coef_var, estimate, `95% CI`, p.value) %>%
  arrange(predict_spec, coef_var)%>%mutate(
    coef_var = case_when(
      coef_var == "coef_sample_T" ~ "T0",
      coef_var == "coef_temp_lag1" ~ "T1",
      coef_var == "coef_temp_lag2" ~ "T2",
      coef_var == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))


cor_table %>%
  gt() %>%
  tab_header(
    title = "Correlation between Slope and Temperature Coefficients"
  ) %>%
  cols_label(
    predict_spec = "Species",
    coef_var = "Time lag",
    estimate = "Correlation (r)",
    `95% CI` = "95% Confidence Interval",
    p.value = "p-value"
  ) %>%
  fmt_markdown(columns = everything())
```
**Table 1**: Species level correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope)



```{r species cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}

cor_results <- df_reg %>%
  pivot_longer(cols = starts_with("coef_"), 
               names_to = "coef_var", values_to = "coef_val") %>%
  group_by(predict_spec, coef_var) %>%
  summarise(
    cor_res = list(cor.test(slope, coef_val) %>% tidy(conf.int = TRUE)),
    .groups = "drop"
  ) %>%
  unnest(cor_res) %>%
  dplyr::select(predict_spec, coef_var, estimate, conf.low, conf.high, p.value)%>%mutate(
    coef_var = case_when(
      coef_var == "coef_sample_T" ~ "T0",
      coef_var == "coef_temp_lag1" ~ "T1",
      coef_var == "coef_temp_lag2" ~ "T2",
      coef_var == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))


ggplot(cor_results, aes(x = coef_var, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ predict_spec,scale="free") +
  theme_bw() +
  labs(
    x = "Temperature variable",
    y = "Correlation"
  ) 


```
**Figure 3**: Species level correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope).




```{r species slopes, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=6}
ggplot(data=df_slopes,aes(x=slope,y=predict_spec,color=predict_spec))+
  geom_point(size=4, alpha=.7)+
  theme_bw()+
  labs(y="species",color="species")
```
**Figure 4**: Distributions of species responses. 

## Window analysis

We looked at how the responsiveness to temperature changed over time using a moving window approach (Fig. 5). We found that responsiveness was highest in the beginning of the experiment. 


```{r}

unique_days <- sort(unique(df_T$day))
unique_days<-unique_days[unique_days<=30]
window_size <- 6
n_windows <- length(unique_days) - window_size + 1


window_list <- lapply(1:n_windows, function(i) {
  days_window <- unique_days[i:(i + window_size - 1)]
  window_df<-df_T[df_T$day %in% days_window, ]
  window_df%>%mutate(window=unique_days[i])
})




window_temp_list<-lapply(window_list,function(w){
  
  df_S<-split(w,w$species_ID)
  
  list_reg<-lapply(df_S,function(s){
    # Filter data
    s_filtered <- s %>% dplyr::filter(is_after_first_zero_day == F)
    
  
    
    # Skip if not enough data to fit model
    if (nrow(s_filtered) < 3) {
      return(NULL)
    }
  
    # Fit models
    lm0 <- lm((biomass) ~ sample_T, data = s_filtered)
    lm1 <- lm((biomass) ~ mean_T_lag1, data = s_filtered)
    lm2 <- lm((biomass) ~ mean_T_lag2, data = s_filtered)
    lm3 <- lm((biomass) ~ mean_T_lag3, data = s_filtered)
   
    # Get species responses
    slope_select<-df_slopes%>%dplyr::filter(nutrients==unique(s$nutrients),temperature==unique(s$temperature),predict_spec==unique(s$predict_spec))
    

  
  
   
    
    data.frame(
      species_ID = unique(s$species_ID),
       coef_sample_T = coef(lm0)[["sample_T"]],
    coef_temp_lag1 = coef(lm1)[["mean_T_lag1"]],
    coef_temp_lag2 = coef(lm2)[["mean_T_lag2"]],
    coef_temp_lag3 = coef(lm3)[["mean_T_lag3"]],
      slope=(slope_select$slope),
      temperature=unique(s$temperature),
      nutrients=unique(s$nutrients),
      predict_spec=unique(s$predict_spec),
      richness=nchar(unique(s_filtered$composition)),
      last_non_zero_day=unique(s$last_non_zero_day),
      window=unique(w$window)
      )
    
  })
  df_reg<-do.call(rbind,list_reg)

})

window_temp_reg<-do.call(rbind,window_temp_list)




window_temp_reg <- window_temp_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2" ~ "T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))
  


window_lag_list<-split(window_temp_reg,window_temp_reg$temp_lag)



List_window_lag<-lapply(window_lag_list,function(lag){
  
      window_temp_list<-split(lag,lag$window)

      list_window_cor<-lapply(window_temp_list,function(w){
        
        
      
          
         result <- cor.test(w$slope, w$coef_value)
         broom::tidy(result) %>%mutate(window=unique(w$window))%>%
         select(window, estimate, conf.low, conf.high, p.value)
         
      
      })
      df_window<-do.call(rbind,list_window_cor)
      df_window<-df_window%>%mutate(temp_lag=unique(lag$temp_lag))
})



df_cor_window<-do.call(rbind,List_window_lag)

  
plot_overal_cor<-ggplot(data = df_cor_window ,
       aes(x = window%>%as.factor, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +

  geom_hline(yintercept=0,linetype="dashed")+
  theme_bw() +
  labs( x = "Start window (day)",y="Correlation")+
  facet_wrap(~temp_lag)





```





```{r window_cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_overal_cor

```
**Figure 5**: Window analysis for T0 - T3: Correlation  between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope). Every window contains 6 sampling days with +1 sampling day displacement between windows. 















