---
title: "Temperature responses: The balance of nature: Critical Role of Species Environmental Responses for Stability"
author: "Til H채mmig, Francesco Polazzo, Owen L. Petchey, Frank Pennekamp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
    keep_md: no
    fig_caption: false  
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
rm(list=ls())
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("here")
library(flextable)
library(performance)
library(lavaan)
library(officer)
library(gt)
library(ggbeeswarm)
library(lmerTest)
library(codyn)
library(broom)

#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")

```

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```



# Load datasets, Data wrangling and Imbalance calculation

```{r  warning=FALSE, results='hide'}

divergence_df <- read_csv(here("Data", "divergence_df.csv"))
load(here("Data", "dens_biomass_poly.RData"))

dd_all_pred<-read.csv(here("Data", "morph_dd_pred.csv"))
dd_all_pred_nonoise<-read.csv(here("Data", "morph_dd_pred_nonoise.csv"))

load(here("Data", "ciliate_traits.Rdata"))

df_slopes <- read_csv(here("Data", "df_slopes_cor.csv"))

# needs to have id_new variable
ciliate_traits <- ciliate_traits %>%
  dplyr::mutate(
    # Remove dots from the date
    cleaned_date = gsub("\\.", "", date),
    # Extract the part of id after the underscore
    id_suffix = sub(".*_(.*)", "\\1", id),
    # Combine cleaned_date, id_suffix, and species_initial into a new variable
    id_new = paste0(cleaned_date, id_suffix, composition)
  ) %>%
  # Optionally, remove the intermediate columns to clean up
  dplyr::select(-cleaned_date, -id_suffix,-new_id)

uniqueN(ciliate_traits$id_new)==nrow(ciliate_traits) # all unique  ;)

id_dd<-full_join(dd_all_pred,dplyr::select(ciliate_traits,id_new,biomass),join_by("id_new"))


## add day variable

#create a day variable from the date variable

id_dd<-dplyr::mutate(id_dd,date=as.Date(date,format = "%d.%m.%y"))

earliest_date<-min(id_dd$date)
days_since_earliest<-as.numeric(id_dd$date-earliest_date)+1
id_dd<-id_dd%>%dplyr::mutate(day=days_since_earliest)

#create a summarised df on microcosm level with each species seperate
# Make sure, that we have n_frames and not N_frames
names(id_dd)[names(id_dd) == "N_frames"] <- "n_frames"

#extrapolation_factor <- 9.301902  # for 16 x magnification 
extrapolation_factor <- 9.828125  # for 25 x magnification 
video_biomass_species <- c( "C", "P", "S","D","L","T")

biomasses <- id_dd %>%
  dplyr::group_by( day,temperature,nutrients,sample_ID,composition,predict_spec) %>% # group  by xxx
  dplyr::summarize(
    biomass = sum(biomass * n_frames, na.rm = TRUE) / (1 * 125) # if not 3 videos corrections is done below with dens_factor
  ) %>%
  dplyr::mutate(
    biomass = biomass * extrapolation_factor,
    )

biomasses<-biomasses%>%dplyr::mutate(biomass=biomass*1000)

dd_ts_id<-biomasses

#fill up missing dates with biomass<-0

fill_dd<-expand.grid(sample_ID=unique(dd_ts_id$sample_ID),day=unique(dd_ts_id$day),predict_spec=unique(dd_ts_id$predict_spec))
complete_ts<-full_join(fill_dd,dd_ts_id,join_by(sample_ID,day,predict_spec))
#complete_ts<-complete_ts%>%dplyr::filter(day>=10)

complete_ts$biomass[is.na(complete_ts$biomass)]<-0
complete_ts<-complete_ts%>%dplyr::mutate(composition=sub("_.*", "", sample_ID))
complete_ts<-complete_ts %>%
  dplyr::mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
complete_ts<- dplyr::mutate(complete_ts,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))

# Now remove wrong combinations of composition and predict_spec / predict_spec

complete_ts<- complete_ts %>%
  rowwise() %>%
  dplyr::filter(predict_spec %in% unlist(strsplit(composition, ""))) %>%
  ungroup()  
complete_ts<-dplyr::mutate(complete_ts,temperature=as.character(temperature),
                    nutrients=as.character(nutrients),
                    richness=nchar(composition))

complete_ts<-complete_ts%>%group_by(sample_ID,composition,day)%>%dplyr::mutate(tot_biomass=sum(biomass))
complete_ts<-complete_ts%>%dplyr::mutate(biom_contribution=biomass/tot_biomass)

df_biomass_mod <- complete_ts

complete_ts<-complete_ts%>%dplyr::mutate(temperature=paste0(temperature," 째C"),
                                      nutrients=paste0(nutrients," g/L"))


# introduce slopes of 
names(df_slopes)[names(df_slopes)=="species_initial"]<-"predict_spec"

slope_ts<-full_join(dplyr::select(df_slopes,nutrients,predict_spec,temperature,slope),complete_ts)
slope_ts<-slope_ts%>%dplyr::mutate(w_slope=biom_contribution*slope,
                            sign=sign(slope))

slope_ts<-slope_ts%>%group_by(sample_ID,temperature,nutrients,richness,composition,day,tot_biomass)%>%dplyr::summarize(
  sum_w_slopes=abs(sum(w_slope)),
                   mean_abs_slope=mean(abs(slope)),
  sum_abs_slope=sum(abs(slope)),
  abs_sum_slope=abs(sum(slope)),
  symmetry=abs(sum(sign)))


slope_ts<-slope_ts%>%dplyr::mutate(richness=as.factor(richness))


##create new variable where it checks, where the last observation =0 is; with complete_ts
aggr_ts <- slope_ts %>%
  group_by( sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = tot_biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()

aggr_ts<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))
#aggr_ts<-aggr_ts%>%dplyr::filter(day>=10)

### cut of time series when species extinct
# aggr_ts_no_ext<-aggr_ts%>%dplyr::filter(is_after_first_zero_day==F)

biomass_ts<-aggr_ts%>%dplyr::group_by(day,temperature,nutrients,richness)%>%summarize(tot_biom=(mean(tot_biomass)),se_tot_biom=sd((tot_biomass))/sqrt(as.numeric(length(tot_biomass))))


```


 
## Species responses to temperature

  T0:    Does temperature on sampling day explain biomass?  
  
  T1:   Does temperature on the day before sampling day explain biomass? 
  
  T2:   Does average temperature over the two days before sampling day explain biomass? 
  
  T3:   Does average temperature over the three days before sampling day explain biomass? 



```{r}

## Create temperature time series patterns
species_ts<-complete_ts%>%mutate(species_ID=paste(sample_ID,predict_spec,sep="_"))

pattern <- c(18, 18, 18, 19.5, 21, 21, 21, 19.5)
T1<-rep(pattern, length.out = 60)
T1_sel<-data.frame(sample_T=T1[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T1, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T1, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T1, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))


pattern <- c(22, 22, 22, 23.5, 25, 25, 25, 23.5)
T2<-rep(pattern, length.out = 60)
T2_sel<-data.frame(sample_T=T2[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T2, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T2, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T2, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))

pattern <- c(25, 25, 25, 26.5, 28, 28, 28, 26.5)
T3<-rep(pattern, length.out = 60)
T3_sel<-data.frame(sample_T=T3[unique(species_ts$day)+1],
                   temp_lag1=dplyr::lag(T3, 1)[unique(species_ts$day)+1],
                   temp_lag2=dplyr::lag(T3, 2)[unique(species_ts$day)+1],
                   temp_lag3=dplyr::lag(T3, 3)[unique(species_ts$day)+1],
                   day=unique(species_ts$day))


## Note that total time series goes from sampling day 1 (second day of the experiment) until the 59th day (here 58)-> 1:58

library(dplyr)
library(slider) # for sliding window operations

# Create the temperature time series for each regime


# Get unique days (+1 if needed)
days <- unique(species_ts$day)+1 

# Helper function to compute sliding means
get_temp_summary <- function(temp_vector, days) {
  tibble(
    day = days-1 ,  # shift back to match original day indexing
    mean_T_lag1 = sapply(days, function(d) if (d >= 2) mean(temp_vector[(d-1)]) else NA)%>%unlist,
    mean_T_lag2 = sapply(days, function(d) if (d >= 3) mean(temp_vector[(d-2):(d-1)]) else NA)%>%unlist,
    mean_T_lag3 = sapply(days, function(d) if (d >= 4) mean(temp_vector[(d-3):(d-2)]) else NA)%>%unlist
  )
}

# Generate summary tables
T1_mean_lag <- get_temp_summary(T1, days)
T2_mean_lag <- get_temp_summary(T2, days)
T3_mean_lag <- get_temp_summary(T3, days)


T1_sel<-full_join(T1_sel,T1_mean_lag)
T2_sel<-full_join(T2_sel,T2_mean_lag)
T3_sel<-full_join(T3_sel,T3_mean_lag)



df_S<-split(species_ts,species_ts$species_ID)
df_S<-mclapply(df_S,function(df){
  if(unique(df$temperature)=="18-21 째C"){
    dfT<-T1_sel
  }
  else if(unique(df$temperature)=="22-25 째C"){
    dfT<-T2_sel
  }
  else{dfT<-T3_sel 
  }
  
  # dfT <- dfT %>%arrange(day)%>%
  # mutate(temp_lag1 = dplyr::lag(T1, 1),
  #        temp_lag2 = dplyr::lag(sample_T, 2))
  
   dfT<-full_join(dfT,df)
   
   dfT<-dfT %>%
  group_by( species_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()
  
   
  
  })
   df_T<-do.call("rbind",df_S)

   



list_reg<-lapply(df_S,function(s){
  # Filter data
  s_filtered <- s %>% dplyr::filter(is_after_first_zero_day == F)
  

  
  # Skip if not enough data to fit model
  if (nrow(s_filtered) < 3) {
    return(NULL)
  }

  # Fit models
  lm0 <- lm((biomass) ~ sample_T, data = s_filtered)
  lm1 <- lm((biomass) ~ mean_T_lag1, data = s_filtered)
  lm2 <- lm((biomass) ~ mean_T_lag2, data = s_filtered)
  lm3 <- lm((biomass) ~ mean_T_lag3, data = s_filtered)
 
  # Get species responses
  slope_select<-df_slopes%>%dplyr::filter(nutrients==unique(s$nutrients),temperature==unique(s$temperature),predict_spec==unique(s$predict_spec))
  
  # Get rest_biomass


 
  
  data.frame(
    species_ID = unique(s$species_ID),
    coef_sample_T = coef(lm0)[["sample_T"]],
    coef_temp_lag1 = coef(lm1)[["mean_T_lag1"]],
    coef_temp_lag2 = coef(lm2)[["mean_T_lag2"]],
    coef_temp_lag3 = coef(lm3)[["mean_T_lag3"]],
    slope=(slope_select$slope),
    temperature=unique(s$temperature),
    nutrients=unique(s$nutrients),
    predict_spec=unique(s$predict_spec),
    mean_biom_cont=mean(s_filtered$biom_contribution,na.rm=T),
    richness=nchar(unique(s_filtered$composition)),
    last_non_zero_day=unique(s$last_non_zero_day)
    )
  
})

df_reg<-do.call(rbind,list_reg)



temp_lags <- c("coef_sample_T", "coef_temp_lag1", "coef_temp_lag2", "coef_temp_lag3")

# Apply cor.test for each temp lag and tidy the results
cor_summary <- map_dfr(temp_lags, function(var) {

  result <- cor.test(df_reg$slope, df_reg[[var]])
  broom::tidy(result) %>%
    mutate(temp_lag = var) %>%
    select(temp_lag, estimate, conf.low, conf.high, p.value)%>%
    mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2"  ~"T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))
})


plot_overal_cor<-ggplot(data = cor_summary ,
       aes(x = temp_lag%>%as.factor, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +

  geom_hline(yintercept=0,linetype="dashed")+
  theme_bw() +
  scale_x_discrete(labels = function(x) parse(text = x)) +
  labs( x = "Temperature variable",y="Correlation")




df_reg_long <- df_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2"  ~"T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))






```



## Overall correlation


```{r}
df_reg_long <- df_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2" ~ "T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))




# Plot with custom facet titles
plot_cor_reg_slope <- ggplot(data = df_reg_long, aes(x = slope, y = coef_value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~temp_lag, nrow = 1) +
  theme_bw()+
  labs(y="Regression coefficient: biomass ~ temperature",x = "Response from monoculture (slope)")
```



```{r overal reg_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_cor_reg_slope
```



**Figure 1**: Relationship between species' responses to temperature from monoculture and from the community experiment. The positive relationship is evidence that species were responding to the fluctuating temperature regime in the community experiment. 

```{r overal cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_overal_cor
```



**Figure 2**: Correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope). Positive correlations indicate that species in polyculture were responding according to the responses estimated from the monocultre experiment. 



## Species level

```{r species cor_table, results='asis', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
# Prepare correlation results
cor_table <- df_reg %>%
  pivot_longer(cols = starts_with("coef_"), names_to = "coef_var", values_to = "coef_val") %>%
  group_by(predict_spec, coef_var) %>%
  summarise(cor_res = list(cor.test(slope, coef_val) %>% tidy(conf.int = TRUE)), .groups = "drop") %>%
  unnest(cor_res) %>%
  select(predict_spec, coef_var, estimate, conf.low, conf.high, p.value) %>%
  mutate(
    estimate = round(estimate, 3),
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3),
    p.value = signif(p.value, 2),
    `95% CI` = paste0("[", conf.low, ", ", conf.high, "]")
  ) %>%
  select(predict_spec, coef_var, estimate, `95% CI`, p.value) %>%
  arrange(predict_spec, coef_var)%>%mutate(
    coef_var = case_when(
      coef_var == "coef_sample_T" ~ "T0",
      coef_var == "coef_temp_lag1" ~ "T1",
      coef_var == "coef_temp_lag2" ~ "T2",
      coef_var == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))

# Create a formatted gt table
cor_table %>%
  gt() %>%
  tab_header(
    title = "Correlation between Slope and Temperature Coefficients"
  ) %>%
  cols_label(
    predict_spec = "Species",
    coef_var = "Time lag",
    estimate = "Correlation (r)",
    `95% CI` = "95% Confidence Interval",
    p.value = "p-value"
  ) %>%
  fmt_markdown(columns = everything())
```
**Table 1**: Species level correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope)



```{r species cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}

cor_results <- df_reg %>%
  pivot_longer(cols = starts_with("coef_"), 
               names_to = "coef_var", values_to = "coef_val") %>%
  group_by(predict_spec, coef_var) %>%
  summarise(
    cor_res = list(cor.test(slope, coef_val) %>% tidy(conf.int = TRUE)),
    .groups = "drop"
  ) %>%
  unnest(cor_res) %>%
  dplyr::select(predict_spec, coef_var, estimate, conf.low, conf.high, p.value)%>%mutate(
    coef_var = case_when(
      coef_var == "coef_sample_T" ~ "T0",
      coef_var == "coef_temp_lag1" ~ "T1",
      coef_var == "coef_temp_lag2" ~ "T2",
      coef_var == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))


ggplot(cor_results, aes(x = coef_var, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ predict_spec,scale="free") +
  theme_bw() +
  labs(
    x = "Temperature variable",
    y = "Correlation"
  ) 


```
**Figure 3**: Species level correlation between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope).



## Window analysis
```{r}


# Assume your data frame is called df and has a column named 'day'
unique_days <- sort(unique(df_T$day))
window_size <- 13
n_windows <- length(unique_days) - window_size + 1

# Create list of data frames using sliding window
window_list <- lapply(1:n_windows, function(i) {
  days_window <- unique_days[i:(i + window_size - 1)]
  window_df<-df_T[df_T$day %in% days_window, ]
  window_df%>%mutate(window=unique_days[i])
})




window_temp_list<-lapply(window_list,function(w){
  
  df_S<-split(w,w$species_ID)
  
  list_reg<-lapply(df_S,function(s){
    # Filter data
    s_filtered <- s %>% dplyr::filter(is_after_first_zero_day == F)
    
  
    
    # Skip if not enough data to fit model
    if (nrow(s_filtered) < 3) {
      return(NULL)
    }
  
    # Fit models
    lm0 <- lm((biomass) ~ sample_T, data = s_filtered)
    lm1 <- lm((biomass) ~ mean_T_lag1, data = s_filtered)
    lm2 <- lm((biomass) ~ mean_T_lag2, data = s_filtered)
    lm3 <- lm((biomass) ~ mean_T_lag3, data = s_filtered)
   
    # Get species responses
    slope_select<-df_slopes%>%dplyr::filter(nutrients==unique(s$nutrients),temperature==unique(s$temperature),predict_spec==unique(s$predict_spec))
    
    # Get rest_biomass
  
  
   
    
    data.frame(
      species_ID = unique(s$species_ID),
       coef_sample_T = coef(lm0)[["sample_T"]],
    coef_temp_lag1 = coef(lm1)[["mean_T_lag1"]],
    coef_temp_lag2 = coef(lm2)[["mean_T_lag2"]],
    coef_temp_lag3 = coef(lm3)[["mean_T_lag3"]],
      slope=(slope_select$slope),
      temperature=unique(s$temperature),
      nutrients=unique(s$nutrients),
      predict_spec=unique(s$predict_spec),
      richness=nchar(unique(s_filtered$composition)),
      last_non_zero_day=unique(s$last_non_zero_day),
      window=unique(w$window)
      )
    
  })
  df_reg<-do.call(rbind,list_reg)

})

window_temp_reg<-do.call(rbind,window_temp_list)




window_temp_reg <- window_temp_reg %>%
  pivot_longer(
    cols = starts_with("coef_"),
    names_to = "temp_lag",
    values_to = "coef_value"
  )%>%mutate(
    temp_lag = case_when(
      temp_lag == "coef_sample_T" ~ "T0",
      temp_lag == "coef_temp_lag1" ~ "T1",
      temp_lag == "coef_temp_lag2" ~ "T2",
      temp_lag == "coef_temp_lag3" ~ "T3",
      TRUE ~ NA_character_))
  
window_temp_reg<-window_temp_reg%>%dplyr::filter(temp_lag=="T1")


window_temp_list<-split(window_temp_reg,window_temp_reg$window)

list_window_cor<-lapply(window_temp_list,function(w){
   result <- cor.test(w$slope, w$coef_value)
   broom::tidy(result) %>%mutate(window=unique(w$window))%>%
    select(window, estimate, conf.low, conf.high, p.value)
  
})

df_cor_window<-do.call(rbind,list_window_cor)

  
plot_overal_cor<-ggplot(data = df_cor_window ,
       aes(x = window%>%as.factor, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +

  geom_hline(yintercept=0,linetype="dashed")+
  theme_bw() +
  labs( x = "Start window (day)",y="Correlation")


# Plot with custom facet titles
plot_cor_window_slope <- ggplot(data = window_temp_reg, aes(x = slope, y = coef_value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw()+
  labs(y="Regression coefficient: biomass ~ temperature",x = "Response from monoculture (slope)")+
  facet_wrap(~window)


```



### Plots windows



```{r window_reg_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_cor_window_slope
```
**Figure 4**: Window analysis: Relationship between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope). Every window contains 13 sampling days with +1 sampling day displacement between windows. 



```{r window_cor_plot, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_overal_cor
```
**Figure 5**: Window analysis on temperature lag = -1 day: Correlation  between regression coefficients (biomass~temperature) from the community experiment and from monoculture responses (slope). Every window contains 13 sampling days with +1 sampling day displacement between windows. 

We can see that after the first two windows, the positive relationship disappears, suggesting that species dynamics are no longer (or only weakly) governed by temperature fluctuations. This observation is in accordance with the window analysis on the effect of imbalance on asynchrony, where we also observe a rapid decline of the relationship after the first two windows. 

But why then is the direct effect of imbalance on community stability robust?

```{r}
ggplot(data=df_T,aes(x=slope,y=last_non_zero_day))+
  geom_boxplot()


# Assume your data frame is called df and has a column named 'day'
unique_days <- sort(unique(df_T$day))
window_size <- 13
n_windows <- length(unique_days) - window_size + 1

# Create list of data frames using sliding window
window_list <- lapply(1:n_windows, function(i) {
  days_window <- unique_days[i:(i + window_size - 1)]
  window_df<-df_T[df_T$day %in% days_window, ]
  window_df%>%mutate(window=unique_days[i])
})

window_acf<-lapply(window_list,function(w){
  
  df_S<-split(w,w$species_ID)
  
  list_reg<-lapply(df_S,function(s){
    # Filter data
    s_filtered <- s#%>% dplyr::filter(is_after_first_zero_day == F)
    
  
 acf_result <- acf(s_filtered$biomass, plot = FALSE)
  slope_select<-df_slopes%>%dplyr::filter(nutrients==unique(s$nutrients),temperature==unique(s$temperature),predict_spec==unique(s$predict_spec))
   
    
    data.frame(
      species_ID = unique(s$species_ID),
      lag1_acf = acf_result$acf[2],
      
      slope=(slope_select$slope),
      temperature=unique(s$temperature),
      nutrients=unique(s$nutrients),
      predict_spec=unique(s$predict_spec),
      richness=nchar(unique(s_filtered$composition)),
      last_non_zero_day=unique(s$last_non_zero_day),
      window=unique(w$window),
      mean_biomass=mean(s_filtered$biomass),
      sd_biomass=sd(s_filtered$biomass),
      sample_ID=unique(s_filtered$sample_ID)
      )
    
  })
  df_reg<-do.call(rbind,list_reg)
  df_reg<-df_reg%>%group_by(sample_ID)%>%
  mutate(imbalance=abs(sum(slope)))
  
  
})



acf_df<-do.call(rbind,window_acf)

auto_cor<-acf_df%>%group_by(window,predict_spec)%>%
  summarise(estimate=t.test(lag1_acf)$estimate,
            upper_ci=t.test(lag1_acf)$conf.int[2],
            lower_ci=t.test(lag1_acf)$conf.int[1],
            mean_biomass=mean(mean_biomass))




plot_auto_cor <- ggplot(data = auto_cor, aes(x = as.factor(window), y = (estimate))) +
  geom_point() +
  geom_errorbar(aes(ymax = upper_ci, ymin = lower_ci), width = 0.2) +


  theme_bw()+
  labs(y="autocorrelation with lag = -1",x = "window")+
  facet_wrap(~predict_spec)


plot_mean_biom_auto<- ggplot(data = acf_df, aes(x = (mean_biomass), y = lag1_acf) )+
  geom_point() +
  geom_smooth(method="lm")+
  theme_bw()+
  labs(y="Autocorrelation with lag = -1",x = "Mean biomass ")+
  facet_wrap(~predict_spec)

plot_mean_biom_auto

plot_mean_biom_auto<- ggplot(data = acf_df, aes(y = log(mean_biomass/sd_biomass), x = log(abs(slope))) )+
  geom_point() +
  geom_smooth(method="lm")+
  theme_bw()+
  labs(y="1/CV",x = "log(abs(slope))")+
  facet_wrap(~window)

plot_mean_biom_auto


```

```{r auto_cor1, results='hide', echo=FALSE, warning=FALSE, fig.align="center", fig.height=5, fig.width=10}
plot_auto_cor

```








```

